Arbitrary Code Injection in Adblock Plus

### 2023-05-02 paper/paper.tex
Update from Overleaf

\usepackage{dirtree}
\usepackage{hyperref}

\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  frame=single,
  breaklines=true,
  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}

@@

\subsection{Adblock Plus business model}
According to Alphabet's financial report, in the last quarter of 2022, the company had a revenue of more than 59,000 million dollars under the category of advertising \cite{alphabet_revenue}. The figures that the online advertising industry manages show why there is a fierce fight between publisher of contents, ad networks and ad-blocking software. Eyeo GmbH business model is not exempt from this dispute. As part of their business model, the company charges fees to big companies to be included in an allow-list as long as the advertisement on their webpages is considered as "acceptable ads". This category is defined by an "acceptable Ads Committee" created by Adblock's company that counts on the presence of several stakeholders such as Amazon and Dell. The acceptable criteria of an ad depends on its placement on a page, its size and its being distinctive from the main content of the page \cite{acceptable_ads}. Although these commercial deals are not public, the Financial Times published that the fee that Adblock Plus charges to the big companies for being in the whitelist is approximately 30\% of the ad revenues archived by unblocked ads on those sites \cite{financial_times}. 
A caveat related to the links between the stakeholders of the ad industry is that since 2022, the new CEO of Eyeo GmbH is Frank Einecke, a former executive of Google for 14 years.

@@

\subsection{The Rewrite Rule: What and Why}
% Author: Chengcheng Ding
In version 3.2 of Adblock Plus, released on July 17, 2018, a new filter option called "rewrite" was introduced. It allows users to rewrite outgoing requests to certain URL(s). In this version, the destination of requests can be rewritten to resources within the same origin. For example, a request to "foo.com/ad.js" can be rewritten to "foo.com/not-ad.js".

@@

\section{Attack Outline}
The threat model can be schematized in an attack outline where each actor involved fulfills the following requirements.
\begin{itemize}
\item Victim website: open redirect combined with fetching and loading JS code provokes worse problems.
\item Victim user: install malicious filter list (may be updated automatically), visit victim website.
\item Attacker: control the malicious filter list.
\item Potential impact.
\end{itemize}

@@

@@

@@

@@

@@

@@

@@

As a response, a Software Engineer working for Adblock Plus stated that "these concerns may represent a risk, a conscious choice we make. One thing we could do is check \$rewrite rules in the known subscriptions (like EasyList) to keep an eye on them" \cite{abp_issue_6622}.

@@

\subsection{Defense by Adblock Plus}
As a mitigation to this security vulnerability, Adblock Plus included a change in its codebase to only allow rewrite redirections to internal resources. These resources are static and included in the source code of the extension. For example "\$rewrite=abp-resource:blank-js" makes a redirection to an empty JavaScript. Other internal resources available are an empty blank text ("abp-resource:blank-text"), a blank CSS ("abp-resource:blank-css") and a transparent gif ("abp-resource:1x1-transparent-gif"), among others. The \$rewrite rule is still present but with a reduced scope, achieving the goal of not only resolving the security concerns of the rewrite but also being able to prevent errors when blocking URLs \cite{abp_filter_guide}.

@@

\section{Conclusion}
% Author: Chengcheng Ding
To conclude this article, we would like to give a review of the vulnerability that allowed arbitrary code execution: The vulnerability is actually composed of multiple low-priority vulnerabilities from different parties. A victim website needs to have open redirect or custom script hosting enabled, which is not even considered a vulnerability for most developers. A victim user needs to apply an Adblock Plus custom filter from another party, which is not expected to lead to an exploit as powerful as arbitrary code injection. Finally, Adblock Plus needs to allow the rewriting of requests to the same origin.
To the parties involved, the corresponding vulnerabilities are most likely considered non-crucial, and will not be prioritized for patching, if patching is considered at all. Unfortunately, all of those little bugs added up to a very formidable exploit. Of course, each party here has good reasons not to prioritize patching those vulnerabilities, as we have studied in the paper. But we believe the major takeaway here is how little vulnerabilities, seemingly innocent on their own, can sum up to a crucial vulnerability, that could potentially affect millions of people. 
\section*{Acknowledgment}
We would like to thank Hanan Hibshi for her insightful comments throughout the course of our research, and for having shown us a prolific field of study in the area of security such as browser extensions. This paper is a final project for the course "Browser Security" of the Master of Science in Information Technology-Information Security (MSIT-IS) and Master of Science in Information Security (MSIS) postgraduate programs at Carnegie Mellon University.

@@

\section{Implementation code}
% Author: Chengcheng Ding

Directory of the victim website:

\dirtree{%
.1 victim.
.2 static.
.3 advertisement.js.
.3 advertisement.png.
.3 {dynamic-replace.js}.
.3 {dynamic.js}.
.3 {replace.png}.
.3 {script.js}.
.3 {static-replace.js}.
.3 {static.js}.
.3 {style.css}.
.3 welcome.png.
.2 templates.
.3 \hyperref[victim-index.html]{index.html}.
.2 {victim.py}.
}

We are selectively showing some of the code for the context of this paper.

\subsection{victim/templates/index.html}
\label{victim-index.html}

\begin{lstlisting}
<!doctype html>

<html>
	<head>
		<title>Victim</title>
		<link rel="stylesheet" href="/static/style.css"></link>
	</head>
	<body>
		<h1>Hello Victim</h1>
		<hr />
		<p>
			Welcome image:
			<img src="/static/welcome.png">
		</p>
		<hr />
		<p>
			Advertisement:
			<span id="ajs" class="border">...</span>
			<span id="ads_png">
				<!-- <img src="/static/advertisement.png"> -->
			</span>
			<script src="/static/advertisement.js"></script>
		</p>
		<hr />
		<p>
			Targeted advertisement:
			<span id="user_info" class="border">...</span>
		</p>
		<hr />
		<p>
			Static JS:
			<span id="sjs" class="border">...</span>
			<script src="/static/static.js"></script>
		</p>
		<hr />
		<p>
			Dynamic JS:
			<span id="djs" class="border">...</span>
			<button id="load">Load</button>
		</p>
		<hr />
		<p>
			Open <a href="/redirect">redirect</a>:
			<input type="name" id="redir_url" value="{{ redir_url }}"
					autocomplete="off"/>
			<br />
			<ul>
				<li class="redir_prefill">/static/dynamic-replace.js</li>
				<li class="redir_prefill">{{ ATTACKER }}/static/exploit.js</li>
				<!-- <li class="redir_prefill">/upload/user.js</li> -->
			</ul>
		</p>
		<hr />
		<p>
			We will host your input at
			<a href="/upload/user.js">/upload/user.js</a>!
			<br />
			<input name="user" id="user" />
		</p>
		<hr />
		<p>
			CSP: <button id="csp_apply">Apply</button>
			<br />
			{% for enable, line in csp_list %}
			<input type="checkbox" id="c{{ loop.index0 }}"
				name="c{{ loop.index0 }}" autocomplete="off"
				{%- if enable %} checked{% endif %}
				{%- if loop.last %} class="last"{% endif %}>
			<label for="c{{ loop.index0 }}"><tt>{{ line }}</tt></label><br>
			{% endfor %}
		</p>
		<script src="/static/script.js"></script>
	</body>
</html>
\end{lstlisting}

Directory of the attacker website

\dirtree{%
.1 attacker.
.2 static.
.3 exploit.js.
.3 script.js.
.3 style.css.
.2 templates.
.3 index.html.
.2 attacker.py.
.2 filter-list.txt.
}

for the full implementation, please find it at our public repository \href{https://github.com/lxylxy123456/14828}{here}

@@

\begin{enumerate}
    \item In the demo when the dynamic javascript is rewritten, the static javascript is also rewritten. If possible, we think separating those scripts in the demo may gave the audience a clearer understanding.
    \item In the "IMPLEMENTATION" section of the draft, it was mentioned that "Adblock Plus version between 3.2 and 3.5.1", yet, the "reproducibility analysis" section mentions that "This attack works on AdBlock Plus 3.2 (released in July 2018) and 3.5.1 (released in April 2019)". We think that rephrasing the later sentence to "This attack works on version between AdBlock Plus 3.2 (released in July 2018) and 3.5.1 (released in April 2019)" might be clearer than before.
    \item As mentioned in the first slides of the final presentation, the business model of Eyeo GmbH is an interesting model. We think it would be worth mentioning about how the business model affects Ablock Plus's usability from a user's perspective.
    \item To our knowledge, the Brave Browser has its own engine for Adblock named AdBlock-rust and it supports Adblock Plus syntax. Is there a possibility that past version of Adblock-rust would also be vulnerable?
    \item On the browser side, Adblock Plus does maintain a version for Microsoft Edge. Would Edge also be vulnerable for the same version of Adblock Plus?
\end{enumerate}

>>>>

\usepackage{hyperref}

\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  frame=single,
  breaklines=true,
  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
  tabsize=4,
}

@@

\subsection{Adblock Plus business model}
According to Alphabet's financial report, in the last quarter of 2022, the company had a revenue of more than 59,000 million dollars under the category of advertising \cite{alphabet_revenue}. The figures that the online advertising industry manages show why there is a fierce fight between publisher of contents, ad networks and ad-blocking software. Eyeo GmbH business model is not exempt from this dispute. As part of their business model, the company charges fees to big companies to be included in an allow-list as long as the advertisement on their webpages is considered as ``acceptable ads''. This category is defined by an ``acceptable Ads Committee'' created by Adblock's company that counts on the presence of several stakeholders such as Amazon and Dell. The acceptable criteria of an ad depends on its placement on a page, its size and its being distinctive from the main content of the page \cite{acceptable_ads}. Although these commercial deals are not public, the Financial Times published that the fee that Adblock Plus charges to the big companies for being in the whitelist is approximately 30\% of the ad revenues archived by unblocked ads on those sites \cite{financial_times}. 
A caveat related to the links between the stakeholders of the ad industry is that since 2022, the new CEO of Eyeo GmbH is Frank Einecke, a former executive of Google for 14 years.

@@

\subsection{The Rewrite Rule: What and Why}
% Author: Chengcheng Ding
In version 3.2 of Adblock Plus, released on July 17, 2018, a new filter option called ``rewrite'' was introduced. It allows users to rewrite outgoing requests to certain URL(s). In this version, the destination of requests can be rewritten to resources within the same origin. For example, a request to ``foo.com/ad.js'' can be rewritten to ``foo.com/not-ad.js''.

@@

\section{Attack Outline}
The attack can be schematized in the fo

\subsection{Victim website}

has an open redirect combined with fetching and loading JS code provokes worse problems.

\subsection{Victim user}

install malicious filter list (may be updated automatically), visit victim website.

\subsection{Attacker}

control the malicious filter list.

\subsection{Potential impact}

@@

index.html:
\begin{lstlisting}[language=HTML]
<p>
    Targeted advertisement:
    <span id="user_info" class="border">...</span>
</p>
\end{lstlisting}

script.js:
\begin{lstlisting}
{
	const req = new XMLHttpRequest();
	const url = new URL('/advertisement', document.baseURI);
	const params = new URLSearchParams();
	params.set('city', 'pittsburgh');
	params.set('role', 'students');
	params.set('company', 'CMU');
	url.search = params.toString();
	req.addEventListener("load", function() {
		document.getElementById("user_info").innerText = req.responseText;
	});
	req.open("GET", url, true);
	req.send();
}
\end{lstlisting}

@@

index.html:
\begin{lstlisting}[language=HTML]
<p>
    Static JS:
    <span id="sjs" class="border">...</span>
    <script src="/static/static.js"></script>
</p>
\end{lstlisting}

static.js:
\begin{lstlisting}
document.getElementById("sjs").innerText = "Static JS loaded";
document.getElementById("sjs").className = "success";
\end{lstlisting}

static-replace.js:
\begin{lstlisting}
document.getElementById("sjs").innerText = "REPLACED JS";
document.getElementById("sjs").className = "warn";
\end{lstlisting}

@@

index.html:
\begin{lstlisting}[language=HTML]
<p>
    Dynamic JS:
    <span id="djs" class="border">...</span>
    <button id="load">Load</button>
</p>
\end{lstlisting}

dynamic.js:
\begin{lstlisting}
document.getElementById("djs").innerText = "Dynamic JS loaded";
document.getElementById("djs").className = "success";
\end{lstlisting}

dynamic-replace.js:
\begin{lstlisting}
document.getElementById("sjs").innerText = "REPLACED JS";
document.getElementById("sjs").className = "warn";
\end{lstlisting}

@@

index.html:
\begin{lstlisting}[language=HTML]
<p>
    Open <a href="/redirect">redirect</a>:
    <input type="name" id="redir_url" value="{{ redir_url }}"
            autocomplete="off"/>
    <br />
    <ul>
        <li class="redir_prefill">/static/dynamic-replace.js</li>
        <li class="redir_prefill">{{ ATTACKER }}/static/exploit.js</li>
        <!-- <li class="redir_prefill">/upload/user.js</li> -->
    </ul>
</p>
\end{lstlisting}

script.js:
\begin{lstlisting}
function send_redir_url() {
	const payload = new FormData();
	payload.append("url", document.getElementById("redir_url").value);
	const req = new XMLHttpRequest();
	req.open("POST", "/redirect");
	req.send(payload);
}

document.getElementById("redir_url").addEventListener("input", send_redir_url);

for (let i of document.getElementsByClassName("redir_prefill")) {
	i.addEventListener("click", function(event) {
		document.getElementById("redir_url").value = this.innerText;
		send_redir_url();
	});
}
\end{lstlisting}

@@

index.html:
\begin{lstlisting}
<p>
    We will host your input at
    <a href="/upload/user.js">/upload/user.js</a>!
    <br />
    <input name="user" id="user" />
</p>
\end{lstlisting}

script.js:
\begin{lstlisting}
document.getElementById("user").addEventListener("input", function(event) {
	const payload = new FormData();
	payload.append("user", this.value);
	const req = new XMLHttpRequest();
	req.open("POST", "/upload/user.js");
	req.send(payload);
});
\end{lstlisting}

@@

index.html:
\begin{lstlisting}
<p>
    CSP: <button id="csp_apply">Apply</button>
    <br />
    {% for enable, line in csp_list %}
    <input type="checkbox" id="c{{ loop.index0 }}"
        name="c{{ loop.index0 }}" autocomplete="off"
        {%- if enable %} checked{% endif %}
        {%- if loop.last %} class="last"{% endif %}>
    <label for="c{{ loop.index0 }}"><tt>{{ line }}</tt></label><br>
    {% endfor %}
</p>
\end{lstlisting}

script.js
\begin{lstlisting}
document.getElementById("csp_apply").addEventListener("click", function(event) {
	event.preventDefault();
	location.reload();
});

function handle_check(event) {
	const payload = new FormData();
	for (let i = 0;; i++) {
		const elem = document.getElementById("c" + i);
		payload.append("c" + i, JSON.stringify(elem.checked));
		if (elem.classList.contains("last")) {
			break;
		}
	}

	const req = new XMLHttpRequest();
	req.open("POST", "/csp-list");
	req.send(payload);
}
\end{lstlisting}

@@

As a response, a Software Engineer working for Adblock Plus stated that ``these concerns may represent a risk, a conscious choice we make. One thing we could do is check \$rewrite rules in the known subscriptions (like EasyList) to keep an eye on them'' \cite{abp_issue_6622}.

@@

\subsection{Defense by Adblock Plus}
As a mitigation to this security vulnerability, Adblock Plus included a change in its codebase to only allow rewrite redirections to internal resources. These resources are static and included in the source code of the extension. For example ``\$rewrite=abp-resource:blank-js'' makes a redirection to an empty JavaScript. Other internal resources available are an empty blank text (``abp-resource:blank-text''), a blank CSS (``abp-resource:blank-css'') and a transparent gif (``abp-resource:1x1-transparent-gif''), among others. The \$rewrite rule is still present but with a reduced scope, achieving the goal of not only resolving the security concerns of the rewrite but also being able to prevent errors when blocking URLs \cite{abp_filter_guide}.

@@

\section{Conclusion}
% Author: Chengcheng Ding
To conclude this article, we would like to give a review of the vulnerability that allowed arbitrary code execution: The vulnerability is actually composed of multiple low-priority vulnerabilities from different parties. A victim website needs to have open redirect or custom script hosting enabled, which is not even considered a vulnerability for most developers. A victim user needs to apply an Adblock Plus custom filter from another party, which is not expected to lead to an exploit as powerful as arbitrary code injection. Finally, Adblock Plus needs to allow the rewriting of requests to the same origin.
To the parties involved, the corresponding vulnerabilities are most likely considered non-crucial, and will not be prioritized for patching, if patching is considered at all. Unfortunately, all of those little bugs added up to a very formidable exploit. Of course, each party here has good reasons not to prioritize patching those vulnerabilities, as we have studied in the paper. But we believe the major takeaway here is how little vulnerabilities, seemingly innocent on their own, can sum up to a crucial vulnerability, that could potentially affect millions of people. 
\section*{Acknowledgment}
We would like to thank Hanan Hibshi for her insightful comments throughout the course of our research, and for having shown us a prolific field of study in the area of security such as browser extensions. This paper is a final project for the course ``Browser Security'' of the Master of Science in Information Technology-Information Security (MSIT-IS) and Master of Science in Information Security (MSIS) postgraduate programs at Carnegie Mellon University.

@@

@@

\begin{enumerate}
    \item In the demo when the dynamic javascript is rewritten, the static javascript is also rewritten. If possible, we think separating those scripts in the demo may gave the audience a clearer understanding.
    \item In the ``IMPLEMENTATION'' section of the draft, it was mentioned that ``Adblock Plus version between 3.2 and 3.5.1'', yet, the ``reproducibility analysis'' section mentions that ``This attack works on AdBlock Plus 3.2 (released in July 2018) and 3.5.1 (released in April 2019)''. We think that rephrasing the later sentence to ``This attack works on version between AdBlock Plus 3.2 (released in July 2018) and 3.5.1 (released in April 2019)'' might be clearer than before.
    \item As mentioned in the first slides of the final presentation, the business model of Eyeo GmbH is an interesting model. We think it would be worth mentioning about how the business model affects Ablock Plus's usability from a user's perspective.
    \item To our knowledge, the Brave Browser has its own engine for Adblock named AdBlock-rust and it supports Adblock Plus syntax. Is there a possibility that past version of Adblock-rust would also be vulnerable?
    \item On the browser side, Adblock Plus does maintain a version for Microsoft Edge. Would Edge also be vulnerable for the same version of Adblock Plus?
\end{enumerate}
